Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-27 sections/eval.tex
Update on Overleaf.

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. SpGEMM focuses on efficiently computing the product of two sparse matrices to produce an output sparse matrix.
The inner-product, outer product, and Gustavson's algorithm are three different approaches to performing SpGEMM. 
Inner-product involves multiplying two matrices by taking dot products of corresponding rows and columns.
Outer-product computes the product of all pairs of elements from the input matrices.
More detailed explanation is available in Section 2.2 of ~\cite{zhang2021gamma}.
Figures ~\ref{fig:spgemm_small} and ~\ref{fig:spgemm_large} show SpGEMM results with Finch and TACO.
Note that inner-product is very inefficient unless sparse matrices are nearly dense because the cost of intersections that incurs a lot of redundant traversals. For the output-product, TACO only supports the dense output: TACO does not support outer product due to its lack of support for scatter into a 2D workspace. So, due to memory constraints, we are unable to run TACO's outer-product for large sparse matrices. On the other hand, 
So, for large sparse matrices, we only compared Gustavson's algorithms.
\changwan{mention TACO's gustavson's algorithm is hard-wired}
As shown in Figure~\ref{fig:spgemm_small}, for inner-product, Finch achieves comparable performance with TACO, and for outer-product, the 
As shown in Figure~\ref{fig:spgemm_large}, for large sparse matrices, Finch outperforms TACO with an average speedup of 1.05.

>>>>

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics, focusing on efficiently computing the product of two sparse matrices to produce an output sparse matrix.

The inner-product, outer product, and Gustavson's algorithm are three different approaches to performing SpGEMM. Inner-product involves multiplying two matrices by taking dot products of corresponding rows and columns, while outer-product computes the product of all pairs of elements from the input matrices. Further explanation is available in Section 2.2 of Zhang et al. [1].

Figures \ref{fig:spgemm_small} and \ref{fig:spgemm_large} illustrate SpGEMM results with Finch and TACO. Notably, inner-product can be very inefficient unless sparse matrices are nearly dense due to the cost of intersections incurring redundant traversals. Additionally, TACO does not support outer product due to its lack of support for scatter into a 2D workspace, limiting its use for large sparse matrices. In contrast, Finch supports scatter into 2D workspace, allowing for comparison with Gustavson's algorithm for large sparse matrices.

As depicted in Figure \ref{fig:spgemm_small}, Finch achieves comparable performance with TACO for inner-product. For large sparse matrices, as shown in Figure \ref{fig:spgemm_large}, Finch outperforms TACO with an average speedup of 1.05. It's worth mentioning that TACO's Gustavson's algorithm is hard-wired.

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. SpGEMM focuses on efficiently computing the product of two sparse matrices to produce an output sparse matrix.
The inner-product, outer product, and Gustavson's algorithm are three different approaches to performing SpGEMM. 
Inner-product involves multiplying two matrices by taking dot products of corresponding rows and columns.
Outer-product computes the product of all pairs of elements from the input matrices.
More detailed explanation is available in Section 2.2 of ~\cite{zhang2021gamma}.
Figures ~\ref{fig:spgemm_small} and ~\ref{fig:spgemm_large} show SpGEMM results with Finch and TACO.
Note that inner-product is very inefficient unless sparse matrices are nearly dense because the cost of intersections that incurs a lot of redundant traversals. For the output-product, TACO only supports the dense output: TACO does not support outer product due to its lack of support for scatter into a 2D workspace. So, due to memory constraints, we are unable to run TACO's outer-product for large sparse matrices. On the other hand, Finch supports scatter into 2D workspace, 
So, for large sparse matrices, we only compared Gustavson's algorithms.
\changwan{mention TACO's gustavson's algorithm is hard-wired}
As shown in Figure~\ref{fig:spgemm_small}, for inner-product, Finch achieves comparable performance with TACO, and for outer-product, the 
As shown in Figure~\ref{fig:spgemm_large}, for large sparse matrices, Finch outperforms TACO with an average speedup of 1.05.
